---
title: "01 Exploratory Data Analysis"
author: "Anders Sundelin"
date: "2024-03-18"
output: html_document
params: 
    cache: "../ownership/.cache"
    output: "../ownership/output"
    reloo: FALSE
    cores: 2
    threads: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(GGally)
library(tidyr)
library(brms)
library(bayesplot)
library(dplyr)
library(tidyverse)
library(ggpubr)
options(dplyr.summarise.inform = FALSE)
```

```{r ingest data}
source("ingest_data.R")
levels(data$repo)
levels(data$committerteam)
levels(data$authorteam)
```

# Exploratory Data Analysis

Introduced duplicates in files, per repo.

```{r}
low_stat <- function(x) quantile(x, 0.25, type=3)
median_stat <- function(x) quantile(x, 0.5, type=3)
high_stat <- function(x) quantile(x, 0.75, type=3)
q95_stat <- function(x) q95(x)
very_high_stat <- function(x) quantile(x, 0.99, type=3)
max_stat <- function(x) max(x)

m1 <- data |> select(ADD, DEL, COMPLEX, DUP, INTROD) |> summarise(across(ADD:INTROD, min)) |> pivot_longer(cols=everything(), values_to="min")
m2 <- data |> select(ADD, DEL, COMPLEX, DUP, INTROD) |> summarise(across(ADD:INTROD, low_stat)) |> pivot_longer(cols=everything(), values_to="q25")
m3 <- data |> select(ADD, DEL, COMPLEX, DUP, INTROD) |> summarise(across(ADD:INTROD, median_stat)) |> pivot_longer(cols=everything(), values_to="median")
m4 <- data |> select(ADD, DEL, COMPLEX, DUP, INTROD) |> summarise(across(ADD:INTROD, high_stat)) |> pivot_longer(cols=everything(), values_to="q75")
m5 <- data |> select(ADD, DEL, COMPLEX, DUP, INTROD) |> summarise(across(ADD:INTROD, q95_stat)) |> pivot_longer(cols=everything(), values_to="q95")
m6 <- data |> select(ADD, DEL, COMPLEX, DUP, INTROD) |> summarise(across(ADD:INTROD, very_high_stat)) |> pivot_longer(cols=everything(), values_to="q99")
m7 <- data |> select(ADD, DEL, COMPLEX, DUP, INTROD) |> summarise(across(ADD:INTROD, max)) |> pivot_longer(cols=everything(), values_to="max")

metrics <- merge(
merge(
  merge(
    merge(
      merge(
        merge(m1, m2, by="name"),
        m3, by="name"),
      m4, by="name"),
    m5, by="name"),
  m6, by="name"),
m7, by="name")
metrics
```

```{r}
data |> filter(INTROD > 0) |> group_by(repo, INTROD) |> tally()
```

Most of the introduced duplicates are small (single-digits), but some are large, ranging into the hundreds for the IntTest repo.

## Clone introduction ratio per team and repository

Proportion of file changes that introduce at least one clone, per team and repository.
Size of the dot is proportional to the number of files changed by the team in the repository.

```{r}
changesPerRepoAndTeam <- data |> group_by(repo, committerteam) |> summarise(fileschanged=n())
zerosPerRepoAndTeam <- data |> filter(INTROD == 0) |> group_by(repo, committerteam) |> summarise(zeros=n())
zeros_ratio <- merge(changesPerRepoAndTeam, zerosPerRepoAndTeam) |> mutate(introdRatio = 1-(zeros/fileschanged)) |> arrange(introdRatio)

(p <- zeros_ratio |> filter(committerteam != "Unknown") |> ggplot(aes(x=committerteam, y=introdRatio, color=committerteam, size=fileschanged)) + geom_point() + facet_wrap(~ repo) + xlab("team") + scale_color_manual(values=COLOR_BY_TEAM) + theme_bw() + scale_y_continuous(limits=c(0,0.27)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
)
```

As we cannot ascertain which team the Unknown team members participate in, we exclude them from this metric.

## Additions and deletions per team and repository

Viewing the (log-transformed) added and removed lines, we see some correlation, and also quite a few "pure additions" and "pure removals" (along the x and y axes, respectively).


```{r}
data |> group_by(repo) |> ggplot(aes(x=logADD, y=logDEL, color=committerteam)) + geom_point() + facet_wrap(~repo)
```

We see that, overall, there are more duplicates present in the IntTest and Jupiter repos (largest one). And there seems to be at least some indications of correlation with complexity (although there are also some duplicates in files where complexity is 0, i.e. the y axis).

```{r}
data |> group_by(repo) |> ggplot(aes(x=logCOMPLEX, y=logDUP)) + geom_point() + facet_wrap( ~ repo) + scale_color_manual(values=COLOR_BY_TEAM) + theme_bw()
```

A more thorough pairs plot reveal that the parameters seem quite independent - at least no obvious correlations are present (though logCOMPLEX and logDUP are somewhat related, as seen above). Plot can be repeated per repo or per team, with the same conclusions.

```{r}
ggpairs(data |> select(logADD, logDEL, logCOMPLEX, logDUP))
```

```{r}
# quantile type 3 to avoid having interpolation of the observations
repo_committerteam <- data |> group_by(repo, committerteam) |> summarize(q95=quantile(INTROD, 0.95, type=3),
                                                                         q99=quantile(INTROD, 0.99, type=3),
                                                                         max=max(INTROD),
                                                                         files=n(),
                                                                         mean_added=mean(ADD),
                                                                         median_added=median(ADD),
                                                                         mean_removed=mean(DEL),
                                                                         median_removed=median(DEL)) |> mutate(team=committerteam)

```


```{r}
repo_committerteam |> ggplot(aes(x=team, y=repo)) +
    geom_tile(aes(fill=files)) +
    geom_text(aes(label=round(files, 0)), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle(paste("Observed number of changes to files by team and repo"))
```

Repositories are named and sized (though not to scale) after the seven neighboring planets, plus the INTTEST repository, which contains integration tests developed in Java.

We note that some teams have not changed files in some repostories, and that some teams have made very many, and some very few, changes to some repositories.

```{r}
repo_committerteam |> ggplot(aes(x=team, y=repo)) +
    geom_tile(aes(fill=mean_added)) +
    geom_text(aes(label=round(mean_added, 0)), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle(paste("Observed mean number of added lines to files by team and repo"))
```
```{r}
repo_committerteam |> ggplot(aes(x=team, y=repo)) +
    geom_tile(aes(fill=median_added)) +
    geom_text(aes(label=round(median_added, 0)), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle(paste("Observed median number of added lines to files by team and repo"))
```

```{r}
repo_committerteam |> ggplot(aes(x=team, y=repo)) +
    geom_tile(aes(fill=mean_removed)) +
    geom_text(aes(label=round(mean_removed, 0)), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle(paste("Observed mean number of removed lines to files by team and repo"))
```
```{r}
repo_committerteam |> ggplot(aes(x=team, y=repo)) +
    geom_tile(aes(fill=median_removed)) +
    geom_text(aes(label=round(median_removed, 0)), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle(paste("Observed median number of removed lines to files by team and repo"))
```

We note that adding lines are more common than removing, and that the Architect team leads in removing lines in most repositories.

## Introduced duplicates in Jupiter

The number of introduced duplicates follow a Poisson-like (or Negative Binomial) distribution, once you exclude all the zeros, for the zero-inflation part.

```{r}
plot_introd_issues_in_repo <- function(aRepo, team_selector) {
  data |> filter(repo == aRepo, team_selector(committerteam)) |> mutate(team = committerteam) |> group_by(team) |> ggplot(aes(x=INTROD, color=team, fill=team)) + geom_histogram(binwidth = 1) + facet_wrap(~ team) + scale_fill_manual(name="team", values=teamcolors) + scale_colour_manual(name="team", values=teamcolors) + ylab("number of changed files") + xlab("Number of duplicates introduced in the change") + ylim(0,65)
}
(p <- plot_introd_issues_in_repo(JUPITER, function(x) x %in% c(RED, BLUE, GREEN, ARCH)) + theme_bw() + theme(legend.position = "bottom"))
```

```{r}
plot_introd_per_team <- function(aRepo, team_selector) {
  data |> filter(repo == aRepo, team_selector(committerteam)) |> mutate(team = committerteam) |> group_by(team) |> ggplot(aes(x=INTROD, color=team, fill=team)) + geom_histogram(aes(y=..density..), binwidth = 1) + facet_wrap(~ team) + ggtitle(paste("Introduced duplicates in repo", aRepo)) + scale_fill_manual(name="team", values=teamcolors) + scale_colour_manual(name="team", values=teamcolors) + ylab("proportion of filechanges") + xlab("Number of duplicates introduced in the change")
}
```

```{r}
(p <- plot_introd_per_team(JUPITER, function(x) x %in% c(ARCH, RED, BLUE, GREEN)) + theme_bw() + scale_y_continuous(limits=c(0,0.04)) )
```


```{r}
(p <- plot_introd_per_team(INTTEST, function(x) x %in% c(ARCH, RED, BLUE, GREEN)) + theme_bw() + scale_y_continuous(limits=c(0,0.04)) )

```
```{r}
(p <- plot_introd_per_team(URANUS, function(x) x %in% c(ARCH, RED, BLUE, GREEN)) + theme_bw() + scale_y_continuous(limits=c(0,0.06)) )
```
```{r}
(p <- plot_introd_per_team(VENUS, function(x) x %in% c(ARCH, RED, BLUE, GREEN)) + theme_bw() + scale_y_continuous(limits=c(0,0.04)) )
```

```{r}
data |> filter(repo == INTTEST, committerteam == ARCH) |> group_by(INTROD) |> tally()
```

```{r}
data |> filter(repo == VENUS, committerteam == ARCH) |> group_by(INTROD) |> tally()
```

```{r}
data |> filter(repo == VENUS, committerteam == GREEN) |> group_by(INTROD) |> tally()
```


```{r}
data |> filter(committerteam %in% c(BLUE,  GREEN), repo == URANUS) |> group_by(committerteam) |> summarize(median(COMPLEX))

```

```{r}
data |> filter(committerteam %in% c(BLUE,  GREEN), repo == MARS) |> group_by(committerteam, INTROD) |> summarize(n=n(), median(COMPLEX)) |> mutate(freq=n/sum(n))

```


```{r}
data |> filter(committerteam %in% c(ARCH, BLUE, RED, GREEN), repo == JUPITER) |> group_by(committerteam, INTROD) |> summarize(n=n()) |> mutate(freq=n/sum(n))
```

There are also differences between teams --- we see that the ARCH team in this repo introduced very few duplicates, whereas the Blue and Green teams were more comparable.

The pattern is even more pronounced in the IntTest repo

```{r}
data |> filter(committerteam %in% c(ARCH, BLUE, RED, GREEN), repo == URANUS) |> group_by(committerteam, INTROD) |> summarize(n=n()) |> mutate(freq=n/sum(n))

```

## Observed distribution of duplicates

```{r}
(p <- data |> filter(committerteam %in% c(ARCH, BLUE, RED, BROWN), repo %in% c(INTTEST, JUPITER, URANUS)) |> group_by(repo, committerteam, INTROD) |> ggplot(aes(x=INTROD, color=committerteam)) + stat_ecdf() + facet_wrap(~ repo) + 
    xlab("Maximum number of introduced duplicates") +
    ggtitle("Observed cumulative distribution of introduced duplicates") +
    scale_color_manual(values=COLOR_BY_TEAM) + theme_bw() + scale_y_continuous(limits = c(0.85,1.0)) 
)

```

```{r}
(p <- data |> filter(committerteam %in% c(ARCH, BLUE, RED, GREEN), repo %in% c(INTTEST, JUPITER, URANUS)) |> group_by(repo, committerteam) |> ggplot(aes(x=COMPLEX, color=committerteam)) + stat_ecdf() + facet_wrap(~ repo) +  scale_x_continuous(trans="log1p", breaks=c(0,1,3,10,50,200, 1000)) +
  scale_color_manual(values=COLOR_BY_TEAM) + theme_bw() + 
  xlab("Existing complexity in changed file (log scale)") +
  ggtitle("Observed cumulative frequency of existing complexity")
)
```

In the Jupiter repo, Team Blue are more likely to make changes in less complex files, relative to the Red team. Same for Arch team.
In the Neptune repo, the Arch team likewise are less likely to change complex files, but the Blue and Red teams are more similar.


```{r}
data |> summarize(max(ADD), max(DEL), max(COMPLEX), max(DUP))
```

```{r}
data |> filter(committerteam %in% c(ARCH, BLUE, RED, GREEN), repo %in% c(JUPITER)) |> group_by(committerteam) |> summarize(n(), max(ADD), max(DEL), max(COMPLEX), max(DUP))
```

```{r}
data |> filter(committerteam %in% c(ARCH, BLUE, RED, GREEN), repo %in% c(JUPITER)) |> group_by(committerteam) |> summarize(n(), q95(ADD), q95(DEL), q95(COMPLEX), q95(DUP))
```


```{r}
(p <- data |> filter(committerteam %in% c(ARCH, BLUE, RED, GREEN), repo %in% c(INTTEST, JUPITER, URANUS)) |> group_by(repo, committerteam) |> ggplot(aes(x=DUP, color=committerteam)) + stat_ecdf() + facet_wrap(~ repo) + scale_x_continuous(trans="log1p", breaks=c(0,1,3,10,50,200, 1000)) +
    scale_color_manual(values=COLOR_BY_TEAM) + theme_bw() + 
  xlab("Duplicates in changed file (log scale)") +
  ggtitle("Observed cumulative frequency of existing duplicates")
)
```

```{r}
(p <- data |> filter(committerteam %in% c(ARCH, BLUE, RED, GREEN), repo %in% c(INTTEST, JUPITER, URANUS)) |> group_by(repo, committerteam) |> ggplot(aes(x=ADD, color=committerteam)) + stat_ecdf() + facet_wrap(~ repo) +
  scale_x_continuous(trans="log1p", breaks=c(0,1,3,10,50,200, 1000)) + #, guide=guide_axis(angle=90)) +
  scale_color_manual(values=COLOR_BY_TEAM) + theme_bw() + 
  xlab("Number of added lines (log scale)") +
  ggtitle("Observed cumulative frequency of added lines")
)
```

```{r}
(p <- data |> filter(committerteam %in% c(ARCH, BLUE, RED, GREEN), repo %in% c(INTTEST, JUPITER, URANUS)) |> group_by(repo, committerteam) |> ggplot(aes(x=DEL, color=committerteam)) + stat_ecdf() + facet_wrap(~ repo) +
  scale_x_continuous(trans="log1p", breaks=c(0,1,3,10,50,200, 1000, 4000)) +
  scale_color_manual(values=COLOR_BY_TEAM) + theme_bw() +
  xlab("Number of deleted lines (log scale)") +
    ggtitle("Observed cumulative frequency of deleted lines") #+ scale_y_continuous(limits = c(0.9,1))
)
```

```{r}
quantile_table <- function(df) {
  df |> mutate(team=committerteam) |> group_by(repo, team) |> summarize(n=n(), q50ADD=median(ADD), q50DEL=median(DEL), q50COMP=median(COMPLEX), q50DUP=median(DUP), q75ADD=quantile(ADD, 0.75, type=3), q75DEL=quantile(DEL, 0.75, type=3), q75COMP=quantile(COMPLEX, 0.75, type=3), q75DUP=quantile(DUP, 0.75, type=3))
}
quantile_table(data |> filter(committerteam %in% c(ARCH, BLUE, RED, GREEN), repo %in% c(INTTEST, JUPITER, URANUS)))
```

```{r}
data |> filter(committerteam %in% c(ARCH, BLUE, RED), repo == JUPITER, DEL <= 3) |> group_by(committerteam) |> tally()
```


```{r}
data |> filter(committerteam %in% c(ARCH, BLUE, RED, GREEN), repo %in% c(INTTEST, JUPITER, URANUS)) |> group_by(repo, committerteam, COMPLEX, logCOMPLEX) |> ggplot(aes(x=COMPLEX, color=committerteam)) + stat_ecdf() + facet_wrap(~ repo) + 
    scale_color_manual(values=COLOR_BY_TEAM) + theme_bw() + scale_x_continuous(trans="log1p", breaks=c(0,3,10,50,200, 1000))
```

```{r}
(p <- plot_introd_issues_in_repo(INTTEST, three_teams) + ylim(0,0.045) )
```

There seem to be some team-level variation, as well as repo-level variation in the number of introduced duplicates.

Comparing the largest repo, and the three largest contributors to that repo reveals that the Blue team is more likely to introduce a small amount of duplicates (it has almost double the amount of single-added duplicates as the Red team, which has a similar amount of contributions). But the red team has some more occurrences of 4-8 duplicates added to a single file.

```{r}
(p <- plot_introd_issues_in_repo(JUPITER, function(x) x %in% c(RED, GREEN,BLUE)) + ylim(0,0.045) )
```

```{r}
( data |> filter(committerteam == BLUE, ISTEST==F) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.04) + ggtitle(paste0("Introduced duplicates in non-test code by team ", BLUE)) )
```

```{r}
(data |> filter(committerteam == BLUE, ISTEST==T) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.04) + ggtitle(paste0("Introduced duplicates in test code by team ", BLUE)) )
```
```{r}
(data |> filter(committerteam == BLUE) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + facet_wrap(~ ISTEST) + scale_y_continuous(labels=scales::percent) + ylim(0,0.025) + ggtitle(paste0("Introduced duplicates overall by team ", BLUE)) )
```
```{r}
(data |> filter(committerteam == BLUE) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.04) + ggtitle(paste0("Introduced duplicates overall by team ", BLUE)) )
```

No particular pattern seen for test code.


```{r}
plot_introd_issues_in_repo(MARS, function(x) x %in% c(RED, GREEN,BLUE))+ ylim(0,0.1)
```
```{r}
plot_introd_issues_in_repo(SATURN, function(x) x %in% c(RED, GREEN,BLUE))+ ylim(0,0.05)
```


Quantile plots

```{r}
repo_committerteam |> ggplot(aes(x=team, y=repo)) +
    geom_tile(aes(fill=q95)) +
    geom_text(aes(label=round(q95, 0)), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle(paste("Observed 95% introductions by team and repo"))
```

The 95% quantile plots show that for many repos, and many teams in those repos, in 19 out of 20 file changes, no duplicates were added. The outlier is INTTEST, where many teams can be expected to introduce single-digit duplicates. In the NEPTUNE repository, we also can expect some more duplicate introduction than in the others.


```{r}
repo_committerteam |> ggplot(aes(x=team, y=repo)) +
    geom_tile(aes(fill=q99)) +
    geom_text(aes(label=round(q99, 0)), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle(paste("Observed 99% introductions by team and repo"))

```

The pattern repeats for the 99% quantile. Integration test can be expected to have more duplicates introduced, and in particular the Brown team stands out.
In the Neptune repo, however, it is the Blue and Green teams that are a bit more likely to introduce duplicates.

We also note that the Architect team are unlikely to introduce duplicates.


```{r}
repo_committerteam|> ggplot(aes(x=team, y=repo)) +
    geom_tile(aes(fill=max)) +
    geom_text(aes(label=round(max, 0)), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle(paste("Observed max number of introduced duplicates by team and repo"))
```
```{r}
repo_committerteam_summary <- repo_committerteam |> mutate(files=replace_na(files, 0)) |> group_by(repo) |> summarize(mx=max(files), mn=min(files))
(p <- repo_committerteam |> inner_join(repo_committerteam_summary) |> mutate(p=(files - mn)/(mx-mn)) |> 
  ggplot(aes(x=team, y=repo)) +
    geom_tile(aes(fill=p)) +
    geom_text(aes(label=round(files, 0)), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle("Observed number of changes to files by team and repo", "Heatmap colored by row")
)
```

The Blue team are the most frequent in changing files, with Red and Green following.
The Pink, UI and Unknown teams are much more distant.

```{r}
repo_committerteam_summary <- repo_committerteam |> mutate(max=replace_na(max, 0)) |> group_by(repo) |> summarize(mx=max(max), mn=min(max))
(p <- repo_committerteam |> inner_join(repo_committerteam_summary) |> mutate(p=(max - mn)/(mx-mn)) |> 
  ggplot(aes(x=team, y=repo)) +
    geom_tile(aes(fill=p)) +
    geom_text(aes(label=round(max, 0)), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle("Observed max number of introduced duplicates by team and repo", "Heatmap colored by row")
)
```

The maximum number of introduced duplicates are more spread out between the teams - Blue, Red and Green all have the lead in one or two repositories.
But Brown leads in two (including Integration tests, with the overall maximum of 150 introduced clones), and Yellow in Mercury, the smallest repository in terms of LOC.


```{r}
repo_committerteam_summary <- repo_committerteam |> mutate(max=replace_na(q99, 0)) |> group_by(repo) |> summarize(mx=max(q99), mn=min(q99))
(p <- repo_committerteam |> inner_join(repo_committerteam_summary) |> mutate(p=(q99 - mn)/(mx-mn)) |> 
  ggplot(aes(x=team, y=repo)) +
    geom_tile(aes(fill=p)) +
    geom_text(aes(label=round(q99, 0)), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle("Observed Q99 number of introduced duplicates by team and repo", "Heatmap colored by row")
)
```

The 99th percentile is more spread out - but Brown still have a commanding lead in the Integration tests, where they introduced 103 duplicates in a single change.


```{r}
data |> group_by(repo, committerteam) |> select(committer) |> distinct() |> tally() |> ggplot(aes(x=committerteam, y=repo)) +
    geom_tile(aes(fill=n)) +
    geom_text(aes(label=n), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle(paste("Observed number of committers by team and repo"))
```
```{r}
data |> group_by(repo, authorteam) |> select(author) |> distinct() |> tally() |> ggplot(aes(x=authorteam, y=repo)) +
    geom_tile(aes(fill=n)) +
    geom_text(aes(label=n), color="white") +
    xlab("team") + ylab("repo") +
    ggtitle(paste("Observed number of authors by team and repo"))
```

The maximum number of introduced duplicates reveals that the Brown team, which also had high 95% value in the IntTest repo, also introduced 150 duplicates, the overall max value, to a single file in the integration test repository. 
Overall, we see the pattern repeat:

* Integration tests are highly likely to involve additions of duplicates
* The architect team overall is highly unlikely to introduce any duplicates
* There seems to be differences between teams in the numer of introduced duplicates, relative to their contributions. Differences between Red and Blue team, and also the Brown team stands out for introducing many duplicates in at least the integration test repository.

```{r}
data |> filter(INTROD >0) |> ggplot(aes(x=D, y=INTROD, color=C)) + geom_point()
```

```{r}
data |> filter(INTROD >0) |> ggplot(aes(x=C, y=INTROD, color=D)) + geom_point()

```

Those plots show that complexity, rather than number of existing duplicates, are more linearly related to number of introduced duplicates.
At the same time, complexity and existing duplicates are also related - it is just that even if DUP is zero, there is no strong evidence that INTROD is low or zero.

```{r}
data |> filter(INTROD>0) |> ggplot(aes(x=C, y=D, color=INTROD)) + geom_point()

```

## Distribution of parameters

### Added lines

```{r}
data |> ggplot(aes(x=repo, y=(1+ADD))) + geom_violin() + geom_boxplot(width=0.1) + scale_y_continuous(trans='log', breaks=c(5, 20, 100, 500, 1000, 3000)) + ylab("Added lines")
```

Overall, at the median, every repository have less than 10 added lines. Some outliers occur.

```{r}
data |> ggplot(aes(x=committerteam, y=(1+ADD))) + geom_violin() + geom_boxplot(width=0.1) + scale_y_continuous(trans='log', breaks=c(5, 20, 100, 500, 1000, 3000)) + ylab("Added lines")
```

Between the teams (regardless of repository), there are more variation.

### Deleted lines

```{r}
data |> ggplot(aes(x=committerteam, y=(1+DEL))) + geom_violin() + geom_boxplot(width=0.1) + scale_y_continuous(trans='log', breaks=c(5, 20, 100, 500, 1000, 3000)) + ylab("Deleted lines")
```

### Existing clones
```{r}
data |> ggplot(aes(x=committerteam, y=(1+DUP))) + geom_violin() + geom_boxplot(width=0.1) + scale_y_continuous(trans='log', breaks=c(5, 20, 100, 500)) + ylab("Existing duplicates")
```

### Changed file complexity

```{r}
data |> ggplot(aes(x=committerteam, y=(1+COMPLEX))) + geom_violin() + geom_boxplot(width=0.1) + scale_y_continuous(trans='log', breaks=c(5, 20, 100, 500, 1000)) + ylab("McCabe complexity")
```

## OCAM metrics 

Data definition:

 * churn: is the per-file-change max value of added and removed lines. So a file change with 2 added and 20 removed lines will have a churn value of 20.
 * addcomplex: is the per-file-added McCabe complexity, or 0 in case the complexity of the file decreased due to the change.
 * delcomplex: is the per-file-deleted McCabe complexity, or 0 in case the complexity of the file increased due to the change.
 * commits: the number of unique commits made by the team
 
 The above definition means that, at minimum, one of either addcomplex or delcomplex is zero, for any given file change. 
 The only way for the churn value to be zero is if an empty file (containing no lines) was removed. This happens four times in the data set.
 
```{r}
ocam_data <- data


(p <- ocam_rank_repo(ocam_metrics(ocam_data) |> dplyr::filter(repo == INTTEST)) + theme(legend.position = "none"))

```
 
 
```{r}
ocam_data <- data
(p <- ocam_rank_repo(ocam_metrics(ocam_data) |> filter(repo == INTTEST)) + theme_bw() )

(p <- ocam_rank_repo(ocam_metrics(ocam_data) |> filter(repo == JUPITER)) + theme_bw() )

(p <- ocam_rank_repo(ocam_metrics(ocam_data) |> filter(repo == SATURN)) + theme_bw() )

(p <- ocam_rank_repo(ocam_metrics(ocam_data) |> filter(repo == URANUS)) + theme_bw() )

ocam_rank_repo(ocam_metrics(ocam_data) |> filter(repo == NEPTUNE)) + theme_bw()
ocam_rank_repo(ocam_metrics(ocam_data) |> filter(repo == VENUS)) + theme_bw()
(p <- ocam_rank_repo(ocam_metrics(ocam_data) |> filter(repo == MARS))  + theme_bw())

ocam_rank_repo(ocam_metrics(ocam_data) |> filter(repo == MERCURY)) + theme_bw()
```

```{r}
myrank <- function(df) {
  df |> pivot_longer(cols=c("Grankcommits", "Frankchurn", "Erankaddcomplex", "Drankdelcomplex"), values_to = "metric") |> 
    group_by(repo) |> ggplot(aes(x=committerteam, y=name)) + geom_tile(aes(fill=metric)) + facet_wrap(~ repo, ncol=1) + 
    geom_text(aes(label=metric)) + xlab("team") + ylab("metric") + coord_fixed() + scale_y_discrete(labels=ocam_labels_plain) + 
    scale_fill_gradient2(name="team rank", guide=guide_legend(reverse=T), midpoint=6, low = "#91BFDB", mid = "#FFFFBF", high = "#FC8D59")  + 
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.25)) + theme(legend.position = "none")
}

(ocam_jupiter <- myrank(ocam_metrics(ocam_data) |> filter(repo %in% c(JUPITER, URANUS))) + theme_bw() + theme(legend.position = "none") + ylab(NULL))
```


```{r}
data |> filter(addedComplexity == 0, removedComplexity > 0)
```

## Conditional independencies

We can test whether our data supports the conditional independencies implied by our DAG.

REM _||_ COMP | TEAM
REM _||_ DUP  | TEAM

These would seem to imply that "knowing TEAM, makes REM and COMPLEX" independent (similar for DUP). Do our data support that claim?

### Is REMOVED independent of COMPLEXITY, once TEAM is considered?

```{r}
d <- data |> select(y=INTROD,
                    A=A,
                    C=C,
                    D=D,
                    R=R,
                    team=committerteam,
                    repo=repo)
formula <- bf(C ~ R + team)
get_prior(data=d,
          family=gaussian,
          formula=formula)
```
```{r, eval=FALSE}
priors <- c(prior(normal(0, 0.5), class = Intercept),
            prior(normal(0, 0.5), class = b),
            prior(exponential(1), class = sigma)
            )

(v <- validate_prior(prior=priors,
               formula=formula,
               data=d,
               family=gaussian)
)
```

```{r cond_ind_R_C}
M_cond_ind <-
  brm(data = d,
      family = gaussian,
      file = cachefile("added-M_cond_ind_R_C"),
      formula = formula,
      prior = priors,
      warmup = 1000,
      iter  = ITERATIONS,
      chains = CHAINS,
      cores = CORES,
      backend="cmdstanr",
      file_refit = "on_change",
      threads = threading(THREADS),
      save_pars = SAVE_PARS,
      opencl = c(0,0),
      adapt_delta = ADAPT_DELTA)
```

```{r}
summary(M_cond_ind)
```

formula <- bf(C ~ R + team)
Population-Level Effects: 
            Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept      -0.13      0.02    -0.17    -0.08 1.00     2644     3897
R              -0.01      0.01    -0.02     0.00 1.00    14359     8763
teamBlue        0.04      0.03    -0.01     0.09 1.00     2819     4884
teamBrown      -0.18      0.03    -0.24    -0.12 1.00     3472     5368
teamGreen       0.19      0.03     0.13     0.24 1.00     3136     5141
teamOrange      0.23      0.03     0.16     0.29 1.00     3481     6153
teamPink        0.26      0.05     0.17     0.36 1.00     6156     7276
teamRed         0.31      0.03     0.25     0.36 1.00     3099     4603
teamUI          0.64      0.12     0.41     0.88 1.00    10512     8039
teamUnknown    -0.19      0.05    -0.28    -0.10 1.00     5352     6718
teamViolet      0.51      0.04     0.44     0.58 1.00     4159     7121
teamYellow      0.08      0.03     0.02     0.14 1.00     3644     5944

We are only really interested in population-level effects of R on C here. And this clearly shows that 0 is within the likely betas (-0.02 to 0)

### Is REMOVED independent of DUPLICATES, once TEAM is considered?

```{r}
formula <- bf(D ~ R + team)
get_prior(data=d,
          family=gaussian,
          formula=formula)
```

```{r}
priors <- c(prior(normal(0, 0.5), class = Intercept),
            prior(normal(0, 0.5), class = b),
            prior(exponential(1), class = sigma)
            )

(v <- validate_prior(prior=priors,
               formula=formula,
               data=d,
               family=gaussian)
)
```

```{r cond_ind_R_D}
M_cond_ind <-
  brm(data = d,
      family = gaussian,
      file = cachefile("added-M_cond_ind_R_D"),
      formula = formula,
      prior = priors,
      warmup = 1000,
      iter  = ITERATIONS,
      chains = CHAINS,
      cores = CORES,
      backend="cmdstanr",
      file_refit = "on_change",
      threads = threading(THREADS),
      save_pars = SAVE_PARS,
      opencl = c(0,0),
      adapt_delta = ADAPT_DELTA)
```

```{r}
summary(M_cond_ind)
```

There might be some weak correlation between DUP (D) and DEL (R) beta 0.12 and CI between 0.11 and 0.13

# Team Blue

```{r}
(p <- data |> filter(committerteam == BLUE, ISTEST==F) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.03) + ggtitle(paste0("Introduced duplicates in non-test code by team ", BLUE)) )

figsave("observed_nontest_blue.pdf", p)
```

```{r}
(p <- data |> filter(committerteam == BLUE, ISTEST==T) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.03) + ggtitle(paste0("Introduced duplicates in test code by team ", BLUE)) )

figsave("observed_test_blue.pdf", p)
```


```{r}
(p <- data |> filter(committerteam == BLUE) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.03) + ggtitle(paste0("Introduced duplicates in repos by team ", BLUE)) )

figsave("observed_blue.pdf", p)
```

```{r}
data |> filter(committerteam == BLUE) |> group_by(repo, INTROD) |> summarize(count=n()) |> ungroup() |> mutate(rel = count/sum(count)) |> filter(INTROD==0) |> arrange(rel)

(rf <- data |> filter(committerteam == BLUE) |> group_by(repo, INTROD) |> summarize(count=n()) |> ungroup() |> mutate(rel = count/sum(count)) |> filter(INTROD==1) |> arrange(rel) )
sum(rf$rel)

```



```{r}
data |> filter(committerteam == RED) |> group_by(repo, INTROD) |> summarize(count=n()) |> ungroup() |> mutate(rel = count/sum(count)) |> filter(INTROD==0) |> arrange(rel)

(rf <- data |> filter(committerteam == RED) |> group_by(repo, INTROD) |> summarize(count=n()) |> ungroup() |> mutate(rel = count/sum(count)) |> filter(INTROD==1) |> arrange(rel) )
sum(rf$rel)
```

```{r}
(p <- data |> filter(committerteam == RED, ISTEST==F) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.025) + ggtitle(paste0("Introduced duplicates in non-test code by team ", RED)) )
figsave("observed_nontest_red.pdf", p)
```
```{r}
(p <- data |> filter(committerteam == RED, ISTEST==T) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.025) + ggtitle(paste0("Introduced duplicates in test code by team ", RED)) )
figsave("observed_test_red.pdf", p)

```



```{r}
(p <- data |> filter(committerteam == RED) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.025) + ggtitle(paste0("Introduced duplicates in repos by team ", RED)) )
figsave("observed_red.pdf", p)
```


```{r}
data |> filter(committerteam == GREEN) |> group_by(repo, INTROD) |> summarize(count=n()) |> ungroup() |> mutate(rel = count/sum(count)) |> filter(INTROD==0) |> arrange(rel)

(rf <- data |> filter(committerteam == GREEN) |> group_by(repo, INTROD) |> summarize(count=n()) |> ungroup() |> mutate(rel = count/sum(count)) |> filter(INTROD==1) |> arrange(rel) )
sum(rf$rel)
```

### Green

```{r}
(p <- data |> filter(committerteam == GREEN, ISTEST==F) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.05) )
figsave("observed_nontest_green.pdf", p)
```

```{r}
(p <- data |> filter(committerteam == GREEN, ISTEST==T) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.05) )
figsave("observed_test_green.pdf", p)
```


```{r}
(p <- data |> filter(committerteam == GREEN) |> ggplot(aes(x=INTROD, fill=repo)) + geom_histogram(aes(y=after_stat(count/sum(count))), binwidth = 1) + scale_y_continuous(labels=scales::percent) + ylim(0,0.05) )
figsave("observed_green.pdf", p)
```

```{r}
data |> filter(repo==MARS) |> group_by(committerteam, INTROD) |> summarize(count=n())
```

## IntTest and Arch

is it true that they rarely introduce new files in IntTest?

```{r}
data |> filter(repo == INTTEST) |> group_by(committerteam, ISNEW) |> summarize(count=n(), pct=100*count/sum(count)) |> mutate(ratio=round(100*count/sum(count), 1))
```

```{r}
(p <- data |> filter(repo == JUPITER) |> group_by(committerteam, DEL) |> ggplot(aes(x=DEL, fill=committerteam)) + geom_histogram(binwidth=30) + facet_wrap(~ committerteam) + scale_y_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(1, 10,100,1000, 5000)) +
ylab("log10(count)") + scale_fill_manual(values=COLOR_BY_TEAM) + theme_bw() + ggtitle("Deleted lines per team", "Jupiter repository") 
)
figsave("observed_deletions_jupiter.pdf", p)
```

```{r}
data |> group_by(committerteam) |> ggplot(aes(x=committerteam, y=logCOMPLEX)) + geom_violin()
```


```{r}
data |> filter(repo == JUPITER) |> ggplot(aes(y=logDEL, x=committerteam, fill=committerteam)) + geom_violin() + scale_fill_manual(values=COLOR_BY_TEAM) + theme_bw()
```

```{r}
data |> filter(repo == JUPITER) |> ggplot(aes(y=A, x=committerteam, colour=committerteam)) + geom_violin() 
```

  Blue=rgb(31,120,180, maxColorValue = 255),
  Brown=rgb(177,89,40, maxColorValue = 255),
  Green=rgb(51,160,44, maxColorValue = 255),
  Orange=rgb(255,127,0, maxColorValue = 255),
  Pink=rgb(251,154,153, maxColorValue = 255),
  QA=rgb(177,177,177, maxColorValue = 255),
  Red=rgb(227,26,28, maxColorValue = 255),

```{r}
(p <- data |> mutate(team=committerteam) |> filter(repo %in% c(JUPITER, URANUS), team %in% c(RED, BLUE)) |> ggplot(aes(y=ADD, x=DEL, size=logCOMPLEX, colour=team, shape=team)) + geom_point() +
#   scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(1, 10,100,1000, 2000)) + scale_y_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(1, 10,100,1000, 2000)) +
   scale_color_manual(values=c(Blue=rgb(31,120,180,40, maxColorValue = 255), Red=rgb(227,26,28,40, maxColorValue = 255))) + scale_alpha(0.1) +  theme_bw() + facet_wrap(~repo)+ ggtitle("Added and changed lines per team and repo, sized by complexity of the file") + theme(legend.position="bottom") )
figsave("observed_add_del_blue_red.pdf", p)
```

```{r}
(p <- data |> mutate(team=committerteam) |> filter(repo %in% c(JUPITER, URANUS), team %in% c(RED, BLUE)) |> ggplot(aes(y=ADD, x=DEL, size=logCOMPLEX, colour=team, shape=team)) + geom_point() +
   scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(1, 10,100,1000, 2000)) + scale_y_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(1, 10,100,1000, 2000)) +
   scale_color_manual(values=c(Blue=rgb(31,120,180,40, maxColorValue = 255), Red=rgb(227,26,28,40, maxColorValue = 255))) +  theme_bw() + facet_wrap(~repo)+ ggtitle("Added and changed lines per team and repo, sized by complexity of the file") + theme(legend.position="bottom") )
figsave("observed_add_del_blue_red.pdf", p)
```

```{r}
(p <- data |> filter(repo %in% c(JUPITER, URANUS, MARS), committerteam %in% c(RED, BLUE, GREEN)) |> ggplot(aes(y=ADD, x=DEL, size=logCOMPLEX, colour=committerteam, shape=committerteam)) + geom_point() +
   scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(1, 10,100,1000, 2000)) + scale_y_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(1, 10,100,1000, 2000)) +
   scale_color_manual(values=c(Blue=rgb(31,120,180,40, maxColorValue = 255), Red=rgb(227,26,28,40, maxColorValue = 255), Green=rgb(51,160,44,40, maxColorValue = 255))) + scale_alpha(0.1) +  theme_bw() + facet_wrap(~repo)+ ggtitle("Added and changed lines per team and repo, sized by complexity of the file") )
```
```{r}
data |> filter(repo == MARS, committerteam %in% c(BLUE, GREEN)) |> ggplot(aes(y=INTROD, x=ADD, colour=committerteam, size=COMPLEX)) + geom_point() +
   scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(1, 10,100,1000, 2000)) +
   scale_color_manual(values=c(Blue=rgb(31,120,180,75, maxColorValue = 255), Red=rgb(227,26,28,40, maxColorValue = 255), Green=rgb(51,160,44,75, maxColorValue = 255))) + 
  theme_bw() + ggtitle("Introduced duplicates per added lines, sized by complexity of the file", "Mars repository") 
```

```{r}
data |> filter(repo == JUPITER, committerteam %in% c(BLUE, RED)) |> ggplot(aes(y=INTROD, x=ADD, colour=committerteam, size=COMPLEX)) + geom_point() +
   scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(1, 10,100,1000, 2000)) +
   scale_color_manual(values=c(Blue=rgb(31,120,180,75, maxColorValue = 255), Red=rgb(227,26,28,40, maxColorValue = 255), Green=rgb(51,160,44,75, maxColorValue = 255))) + 
  theme_bw() + ggtitle("Introduced duplicates per added lines, sized by complexity of the file", "Jupiter repository") 
```

```{r}
data |> filter(repo == INTTEST, committerteam %in% c(BLUE, RED)) |> ggplot(aes(y=INTROD, x=ADD, colour=committerteam, size=COMPLEX)) + geom_point() +
   scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(1, 10,100,1000, 2000)) +
   scale_color_manual(values=c(Blue=rgb(31,120,180,75, maxColorValue = 255), Red=rgb(227,26,28,40, maxColorValue = 255), Green=rgb(51,160,44,75, maxColorValue = 255))) + 
  theme_bw() + ggtitle("Introduced duplicates per added lines, sized by complexity of the file", "Int.test repository") 

```

```{r}
(p <- data |> filter(repo == URANUS, committerteam %in% c(RED, BLUE)) |> ggplot(aes(y=ADD, x=DEL, size=C, colour=committerteam)) + geom_point(position="jitter") + scale_color_manual(values=c(Blue=rgb(31,120,180,50, maxColorValue = 255), Red=rgb(227,26,28,50, maxColorValue = 255))) + scale_alpha(0.1) +  theme_bw() + ggtitle("Added and changed lines per team, sized by complexity of the file", "Uranus repository") )
figsave("observed_add_del_blue_red_uranus.pdf", p)
```
